---
title: "NonLinear Regression - Presentation"
author: "Gabriel Santos / Anjal Hussan"
date: "2023-04-18"
always_allow_html: yes
output:
  html_document:
    df_print: kable
    theme: flatly
    highlight: pygments
    code_folding: hide
    css: ./lab.css
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 2
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dataset:

Cars and salespersons

we would like to analyze the relationship between how many total cars have been sold by each sales person and how many weeks each salesperson has worked to sell this cars.

# Libraries


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
SalesCars <- read_csv('https://raw.githubusercontent.com/GabrielSantos33/Non-linear-regression/main/CarsSoldWeek.csv', show_col_types = FALSE)
```

# Data Set

```{r}
SalesCars
```
## Plotting the data

```{r}
ggplot(SalesCars, aes(x = J_period, y = Cars_sold)) +
  geom_point() +
  labs(title = "Scatter Plot", 
       x = "Job period", 
       y = "Total cars sold")
```


Does it look linear? Does it look non linear?, both?

What model best represents this data

Does this linear regression fits the data quite well


# Linear model

```{r}
ggplot(SalesCars, aes(x = J_period, y = Cars_sold)) +
  geom_point() +
  geom_smooth(method = "lm", formula = "y ~ x",se = FALSE) +
  labs(title = "Scatter Plot Example", 
       x = "Job period", 
       y = "Total cars sold")
```


## Fit linear regression model

```{r}
lm_model <- lm(Cars_sold ~ J_period, data = SalesCars)
```

## Coefficients

```{r}
coeficient <- coef(lm_model)
```

## Extract R-squared value

```{r}
r_squared <- summary(lm_model)$r.squared
```

## Predict y-values

```{r}
predicted_y <- predict(lm_model)
```

## Adding R-squared value and predicted y-values to dataframe

```{r}
SalesCars$r_squared <- r_squared
SalesCars$predicted_y <- predicted_y
```


## Display R-squared value and predicted y-values

```{r}
cat("R-squared value: ", round(r_squared, 2), "\n")
cat("Predicted Y-values: ", paste(round(predicted_y, 2), collapse = ", "))
```

## Extract coefficients of the linear equation

```{r}
intercept <- coef(lm_model)[1]
slope <- coef(lm_model)[2]
# Display the linear equation
cat("Linear equation: y =", round(intercept, 2), "+", round(slope, 2), "* x")
```
For each week a person has add a bit more than a half car per week

For every week the sells is expected to increase by 0.58 per week

NOTE: for rare cases where setting the intercept to ZERO makes sense. but for this case we will not set it this up.

## Linear regression Output

```{r}
print(summary(lm_model))
```
Linear equation: y = 114.5 + 0.58 * x
Multiple R = 0.8955 very high
Adjusted R-squared:  0.7867
R-squared:  0.8019 this model explain the 80% of the variance
standard error: 45.94 how well the observation fit around the regression line

## ANOVA results

```{r}
anova_results <- anova(lm_model)
```

```{r}
print("ANOVA Results:")
print(anova_results)
```
Regression = mean square of 111097 
Residuals = 27440
total sum of squares = 138537
---------
Mean Sq = 111097
Mean Sq error or residual = 2111 = low relative to the total sum of squares
F Statistics = 52.633 very high you can se that our significance value is very small- the model is statically significant
P-value = 6.32e-05
Jperiod = 6.41e-06


## Residuals

Let's see the residuals

```{r}
residuals <- residuals(lm_model)
# Create a residuals plot
plot(residuals, main = "Residuals Plot", xlab = "Predicted Cars_sold", ylab = "Residuals")
#x = J_period, y = Cars_sold)
```


## Transforming the exploratory valiables

Simply by transforming the exploratory valiables does not make a non-linear model

```{r}
ggplot(SalesCars, aes(x = J_period, y = Cars_sold)) +
  geom_point() +
  geom_smooth(method = "lm", formula = "y ~ log(x)",se = FALSE) +
  labs(title = "Non linear Scatter Plot Example - transformation", 
       x = "Job period", 
       y = "Total cars sold")
```

# Exponetial model

Lets use an exponetial model
not a good model

```{r}
ggplot(SalesCars, aes(x = J_period, y = Cars_sold)) +
  geom_point() + 
  
  geom_smooth(method = "nls", formula = y ~ a * exp(b * x), se = FALSE, color = "red", 
              method.args = list(start = list(a = 5, b = 0))) + # Add a red 
  labs(title = "Exponetial model", 
       x = "X", y = "Y") + # Set title and axis labels
  theme_minimal() 
```

# Non-linear models

## Modelsq quadratic regression model


Non Linear model using modelsq quadratic regression model
not a good one


```{r}

library(ggplot2)

y <- SalesCars$Cars_sold
x <- SalesCars$J_period

# Fit non-linear regression model
modelsq <- nls(y ~ a * x^2 + b * x + c, start = list(a = 1, b = 1, c = 1))
           
p <- ggplot(SalesCars, aes(x = J_period, y = Cars_sold)) +
  geom_point() +
  labs(x = "X", y = "Y")

# Add fitted line to the plot
p <- p +
  stat_function(fun = function(x) predict(modelsq, newdata = data.frame(x = x)), 
                color = "red")

print(p)
```


## Non-linear regression Output

```{r}
print(summary(modelsq))
```

## residual Square model
```{r}
residuals <- residuals(modelsq)
# Create a residuals plot
plot(residuals, main = "Residuals Plot", xlab = "Predicted Cars_sold", ylab = "Residuals")
#x = J_period, y = Cars_sold)
```


